---
title: "STAT5014 Homewoek 2"
author: "Yixin Chen"
date: "09/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
```

  
## Problem 3

Version control will help me keep track of any changes I made to the project. It will be needed in the final project of this course. Because there will be multiple people working on a project and we need to review the history of the code during the project.

## Problem 4

a. Sensory data from five operators.   
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat> 

First, import the data.
```{r sensory_import, cache = FALSE}
url1 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensory_raw <- fread(url1,fill = T, skip = 1, data.table = F)
saveRDS(sensory_raw, "sensory.RDS")
sensory_raw <- readRDS("sensory.RDS")
```
Second, clean the data with base R.
```{r sensory_tidy_baseR, cache = FALSE}
sensory_tidy <- data.frame(item = rep(seq(1,10), each=15),
                           operator = rep(c(1:5),30),
                           n = NA)
temp_n<-matrix(NA,15,10)
for (i in 1:10) {
  temp_n[,i] <- as.numeric(c(sensory_raw[3*i-2,-1], sensory_raw[3*i-1,-6],
                             sensory_raw[3*i,-6]))
}
sensory_tidy$n <- as.vector(temp_n)
```

b. Gold Medal performance for Olympic Men's Long Jump, year is coded as 1900=0.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat>  

First, import the data.
```{r longj_import, cache = FALSE}
url2 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
longj_raw <- fread(url2,skip = 1,fill = T)
saveRDS(longj_raw, "longj.RDS")
longj_raw <- readRDS("longj.RDS")
```
Second, clean the data with base R.
```{r longj_tidy_baseR, cache = FALSE}
longj_tidy_br <- data.frame(Year = c(longj_raw$V1, longj_raw$V3, longj_raw$V5, longj_raw$V7),
                         LongJump = c(longj_raw$V2, longj_raw$V4, longj_raw$V6, longj_raw$V8))
longj_tidy_br <- longj_tidy_br[1:22,]
```

Third, clean the data with tidyverse.
```{r longj_tidy_tidyverse, cache = FALSE}
#colnames(longj_raw) <- rep(c("Year","LongJump"),4)
year <- longj_raw %>%
                 select(V1,V3,V5,V7) %>%
                 gather("V",value="Year",1,2,3,4) %>%
                 select(Year) %>%
                 slice(1:22)
longjump <- longj_raw %>%
                 select(V2,V4,V6,V8) %>%
                 gather("V",value="LongJump",1,2,3,4) %>%
                 select(LongJump) %>%
                 slice(1:22)
longj_tidy_tv <- bind_cols(year, longjump)
```

c. Brain weight (g) and body weight (kg) for 62 species.    
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat>  

First, import the data.
```{r bbweight_import, cache = FALSE}
url3 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
bbweight_raw <- fread(url3,skip = 1,fill = T)
saveRDS(bbweight_raw, "bbweight.RDS")
bbweight_raw <- readRDS("bbweight.RDS")
```
Second, clean the data with base R.
```{r bbweight_tidy_baseR, cache = FALSE}
bbweight_tidy_br <- data.frame(BodyWt = c(bbweight_raw$V1, bbweight_raw$V3, bbweight_raw$V5),
                         BrainWt = c(bbweight_raw$V2, bbweight_raw$V4, bbweight_raw$V6))
bbweight_tidy_br <- bbweight_tidy_br[1:62,]
```

Third, clean the data with tidyverse.
```{r bbweight_tidy_tidyverse, cache = FALSE}
#colnames(longj_raw) <- rep(c("Year","LongJump"),4)
BodyWt <- bbweight_raw %>%
                 select(V1,V3,V5) %>%
                 gather("V",value="BodyWt",1,2,3) %>%
                 select(BodyWt) %>%
                 slice(1:62)
BrainWt <- bbweight_raw %>%
                 select(V2,V4,V6) %>%
                 gather("V",value="BrainWt",1,2,3) %>%
                 select(BrainWt) %>%
                 slice(1:62)
bbweight_tidy_tv <- bind_cols(BodyWt, BrainWt)
```

d. Triplicate measurements of tomato yield for two varieties of tomatos at three planting densities.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat>  

First, import the data.
```{r tomato_import, cache = FALSE}
url4 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tomato_raw <- fread(url4,skip = 1)
saveRDS(tomato_raw, "tomato.RDS")
tomato_raw <- readRDS("tomato.RDS")
```

Second, clean the data with base R.
```{r tomato_tidy_baseR, cache = FALSE}
colnames(tomato_raw)<-c("Variety","d10k","d20k","d30k")
d10k<- strsplit(as.character(tomato_raw$d10k),',') 
d20k<- strsplit(as.character(tomato_raw$d20k),',') 
d30k<- strsplit(as.character(tomato_raw$d30k),',') 
Ife<- c(as.numeric(d10k[[1]]), as.numeric(d20k[[1]]),as.numeric(d30k[[1]]))
ped<- c(as.numeric(d10k[[2]]), as.numeric(d20k[[2]]),as.numeric(d30k[[2]]))
tomato_tidy_br <- data.frame(Variety = rep(c("Ife#1","PusaEarlyDwarf"), each = 9),
                             Density = rep(rep(c(10000,20000,30000),each = 3),2),
                             Yield = c(Ife,ped))
```

Third, clean the data with tidyverse.
```{r tomato_tidy_tidyverse, cache = FALSE}
d10k <- tomato_raw %>%
                  separate(d10k, into = c("r1_10k","r2_10k","r3_10k") ,sep = ",") %>%
                  select(Variety,r1_10k,r2_10k,r3_10k) %>%
                  gather(key = "Density", value = "Yield",2:4)

d20k <- tomato_raw %>%
                  separate(d20k, into = c("r1_20k","r2_20k","r3_20k") ,sep = ",") %>%
                  select(Variety,r1_20k,r2_20k,r3_20k) %>%
                  gather(key = "Density", value = "Yield",2:4)

d30k <- tomato_raw %>%
                  separate(d30k, into = c("r1_30k","r2_30k","r3_30k") ,sep = ",") %>%
                  select(Variety,r1_30k,r2_30k,r3_30k) %>%
                  gather(key = "Density", value = "Yield",2:4)
tomato_tidy_tv <- bind_rows(d10k,d20k,d30k)
tomato_tidy_tv$Density <- rep(c(10000,20000,30000), each = 6)
tomato_tidy_tv %>% group_by(Variety)
```
## Problem 5

Finish this homework by pushing your changes to your repo.  In general, your workflow for this should be:  

1. git pull -- to make sure you have the most recent repo  
2. In R: do some work  
3. git add -- this tells git to track new files  
4. git commit -- make message INFORMATIVE and USEFUL  
5. git push -- this pushes your local changes to the repo  

If you have difficulty with steps 1-5, git is not correctly or completely setup.  See me for help.

**Only submit the .Rmd and .pdf solution files.  Names should be formatted HW2_lastname.Rmd and HW2_lastname.pdf**

## Optional preperation for next class:  

TBD



